Appendix 3: Example of a spatially varying effect of climate on citizen science count data with sdmTMB

### 2022-04-12 doc/appendix-snow-sdmTMB.Rmd
Tweak owl appendix

```{r mesh, echo=TRUE, cache=FALSE, out.width="4in"}
mesh <- make_mesh(snow, xy_cols = c("X", "Y"), cutoff = 1.5)

@@

We can also test for zero-inflation using the `simulate.sdmTMB` function and a couple functions from the DHARMa package.
If our data were zero-inflated or overdispersed, we might have needed to use a delta ("hurdle") model instead. 
This is not built directly into sdmTMB, but can be achieved by fitting the two components separately. See the steps in the vignette [Fitting delta models with sdmTMB](https://pbs-assess.github.io/sdmTMB/articles/delta-models.html).

@@

We can then compress the annual predictions into mean values for plotting the spatially varying effect of NAO.
Most simply, we can extract the overall coefficients using the `b <- tidy(m)` and add them together with the spatially varying coefficients `zeta_s`.
However, this method does not give us spatially explicit estimates of uncertainty.
To do this, we can get the output from the joint precision matrix simulation draws by using the `predict()` function again, but this time indicating a number of simulation draws (e.g., `sims = 500`) and what component we want estimated (`sims_var = "zeta_s"`). 
We also extract the same number of draws for estimates of the main effect of NAO using the `spread_sims()` function (this time `n_sims = 500`). 
The simulation draws for both can then be combined before calculating the median and lower and upper quantiles from these draws (in this case we will use 80%) and adding these values to our prediction data frame.

@@

n_sims <- 500L
zeta_sim <- predict(m, sims = n_sims, sims_var = "zeta_s")
sims <- spread_sims(m, n_sims = n_sims)
beta <- sims$nao
combined <- beta + t(zeta_sim)

@@

```{r plot-map, echo=TRUE, eval=TRUE, out.width="90%", cache=FALSE}
ggplot(data = p_proj) +
  geom_point(
    data = p_mean, aes(X, Y, colour = nao_effect, size = mean_est_count), alpha = 0.5
  ) +
  geom_sf(data = coast, colour = "gray50") +
  geom_sf(data = lakes, colour = "gray50", fill = NA) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000), # adjusts space on sides
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(
    limit = c(min(p_mean$nao_lwr), max(p_mean$nao_upr)), 
    guide = guide_colourbar(direction = "horizontal", title.position = "top")
  ) +
  guides(size = "none") +
  labs(colour = "Estimated effect in log space") +
  ggtitle("Map of combined main and spatially varying effects of NAO on Snowy Owl count") +
  theme(legend.position = c(0.1, 0.07), axis.title = element_blank())
```

```{r plot-map-ci, echo=F, eval=TRUE, out.width="90%", cache=FALSE}
p1 <- ggplot(data = p_proj) +
  geom_point(
    data = p_mean,
    aes(X, Y, colour = nao_lwr, size = mean_est_count), alpha = 0.5
  ) +
  geom_sf(data = coast, colour = "gray50") +
  geom_sf(data = lakes, colour = "gray50", fill = NA) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000), # adjusts space on sides
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(
    limit = c(min(p_mean$nao_lwr), max(p_mean$nao_upr)), 
    guide = guide_colourbar(direction = "horizontal", title.position = "top")
  ) +
  guides(size = "none", colour = "none") +
  # labs(colour = "Lower 90% CI") +
  ggtitle("Lower 80% CI") +
  theme(legend.position = c(0.1, 0.1), axis.title = element_blank())

>>>>

```{r mesh, echo=TRUE, out.width="4in"}
mesh <- make_mesh(snow, xy_cols = c("X", "Y"), cutoff = 1.5)

@@

We can also test for zero-inflation using the `simulate.sdmTMB()` function and a couple functions from the DHARMa package.
If our data were zero-inflated or overdispersed, we might have needed to use a delta ("hurdle") model instead. 
This is not built directly into sdmTMB, but can be achieved by fitting the two components separately. See the steps in the vignette [Fitting delta models with sdmTMB](https://pbs-assess.github.io/sdmTMB/articles/delta-models.html).

@@

We can then compress the annual predictions into mean values for plotting the spatially varying effect of NAO.
Most simply, we can extract the overall coefficients using the `b <- tidy(m)` and add them together with the spatially varying coefficients `zeta_s`.
However, this method does not give us spatially explicit estimates of uncertainty.
To do this, we can get the output from the joint precision matrix simulation draws by using the `predict()` function again, but this time indicating a number of simulation draws (e.g., `sims = 500`) and what component we want estimated (`sims_var = "zeta_s"`). 
We also extract the same number of draws for estimates of the main effect of NAO using the `spread_sims()` function (this time `nsim = 500`). 
The simulation draws for both can then be combined before calculating the median and lower and upper quantiles from these draws (in this case we will use 80%) and adding these values to our prediction data frame.

@@

nsim <- 500
zeta_sim <- predict(m, nsim = nsim, sims_var = "zeta_s")
sims <- spread_sims(m, nsim = nsim)
beta <- sims$nao
combined <- beta + t(zeta_sim)

@@

```{r plot-map, echo=TRUE, eval=TRUE, out.width="90%", echo=FALSE, cache=FALSE}
ggplot(data = p_proj) +
  geom_point(
    data = p_mean, aes(X, Y, colour = nao_effect, size = mean_est_count), alpha = 0.5
  ) +
  geom_sf(data = coast, colour = "gray50") +
  geom_sf(data = lakes, colour = "gray50", fill = NA) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000), # adjusts space on sides
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(
    limit = c(min(p_mean$nao_lwr), max(p_mean$nao_upr)), 
    guide = guide_colourbar(direction = "horizontal", title.position = "top")
  ) +
  guides(size = "none") +
  labs(colour = "Estimated effect in log space") +
  ggtitle("Map of combined main and spatially varying effects of NAO on Snowy Owl count") +
  theme(legend.position = c(0.1, 0.07), axis.title = element_blank())
```

```{r plot-map-ci, echo=F, eval=TRUE, out.width="90%", cache=FALSE, fig.asp=0.3}
p1 <- ggplot(data = p_proj) +
  geom_point(
    data = p_mean,
    aes(X, Y, colour = nao_lwr, size = mean_est_count), alpha = 0.5
  ) +
  geom_sf(data = coast, colour = "gray50") +
  geom_sf(data = lakes, colour = "gray50", fill = NA) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000), # adjusts space on sides
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(
    limit = c(min(p_mean$nao_lwr), max(p_mean$nao_upr)), 
    guide = guide_colourbar(direction = "horizontal", title.position = "top")
  ) +
  guides(size = "none", colour = "none") +
  # labs(colour = "Lower 90% CI") +
  ggtitle("Lower 80% CI") +
  theme(legend.position = c(0.1, 0.1), axis.title = element_blank())
